Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-27 experiments.tex
Draft of Comparison Setup subsection

%------------------------------------------------------------------------------
\subsection{Evaluation Flow}

To evaluate and compare the performance indexes of different
architectures for a specific application, one needs to simulate the
architecture running the code of the application. To make
architectural exploration possible both the compiler and the simulator
have to be retargetable. Trimaran~\cite{trimaran_hp} provides these
tools and thus represents the pillar around which we have constructed
\ee~\cite{palpatti_estimedia03}. \ee{} is a framework that not only
allows us to evaluate any instance of a platform in terms of area,
performance and power, but also implements various techniques for
exploration of the design space.

@@

\subsection{Terms of comparison}

The results presented in the next section compare PS with one widely
adopted Multi-objective Genetic Algorithm
(MOGA)~\cite{knowles_techrep06}. To the best of our knowledge, genetic
approaches in general still represent the best compromise between
efficiency (i.e. time required for exploration) and accuracy of the
reported Pareto Sets.  Note that, for the reasons described above,
mono-objective approaches have been explicitly excluded from this
analysis.  We also exclude other approaches (e.g.  dependency
based~\cite{givargis_tvlsi02}) that make use of some a-priori
knowledge about the role and meaning of each parameter, since our aim
is evaluate the capacity of focusing on interesting parameter regions
without any external judgement or help based on heuristics.  Finally,
it should be pointed out that in this introductory work we will
explicitly use the \emph{SPEA}~\cite{zitzler_eurogen01} variant of
MOGA, remanding to works like~\cite{zitzler_ec00}~\cite{zitzler_tec03}
for a detailed analysis on how it differs from other genetic approaches
and what could be expected from similar multi-objective strategies.

@@

\begin{figure}
  \figLC{pareto_fronts_500}{Pareto fronts found by PS and GA for a fixed budget of 500 configurations.\comment{AA: nelle due figure centrali la linea di GA e' continua, mentre in tutte le altre figure e' tratteggiata. Il revisore ci chiede di uniformare. Non so come riottenere le figure. Potresti pensarci tu?}}
  \begin{center}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
  \end{center}
\end{figure}
Low budget Pareto sets are shown in \figR{pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly dependent upon the application being considered.
These results are obtained in about 5 eras and demonstrate that with a so limited budget PS can compete with
genetic approach while even discovering very different Pareto fronts. 
Considering the extended simulation budget as depicted in
\figR{pareto_fronts_500}, we can qualitatively observe that
even if PS does not strictly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc.

>>>>

%------------------------------------------------------------------------------
\subsection{Configuration Evaluation Flow}

To evaluate and compare the performance indexes of different
architectures for a specific application, one needs to simulate the
architecture running the code of the application. To make
architectural exploration possible both the compiler and the simulator
have to be retargetable. The open source project Trimaran~\cite{trimaran_hp} provides these
tools and together with estimation framework \ee~\cite{palpatti_estimedia03} has been
adopted as a tested and flexible platform for carrying out the
experiments described in the previous subsection.

@@

\subsection{Comparison Setup}

The results presented in the next section compare PS with one widely
adopted Multi-objective Genetic Algorithm~\cite{knowles_techrep06}. To
the best of our knowledge, genetic approaches in general still
represent the best compromise between efficiency (i.e. time required
for exploration) and accuracy of the reported Pareto Sets.  Note that,
for the reasons described above, mono-objective approaches have been
explicitly excluded from this analysis.  We also exclude other
approaches (e.g.  dependency based~\cite{givargis_tvlsi02}) that make
use of some a-priori knowledge about the role and meaning of each
parameter, since our aim is evaluate the capacity of focusing on
interesting parameter regions without any external judgement or help
based on heuristics.  In particular in this work when referring to
genetic approach (or GA) we will denote the widely spread 
\emph{SPEA}~\cite{zitzler_eurogen01} variant of the Multi-objective
Genetic Algorithm, remanding to
works like~\cite{zitzler_ec00}~\cite{zitzler_tec03} for a detailed
analysis on how it could differ from other genetic approaches and what
could be expected from similar multi-objetive strategies.

\comment{TODO controllare inglese}
Even if an analogy between the concepts of ``era'' and ``generation'' 
can be considered, a fixed amount of total different simulations has been
considered for each of the tests. The purpose is to make the
comparison fair from a algorithm performance perspective, since a
fixed budget of simulations is directly related to the amount of time
required to execute the exploration; on the other side, simply
referring to number of generations/era could be misleading when a
significative number of configuration remain unaltered in the current
population (e.g. because of a local optimum).

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, an accurate
investigation of algorithm tuning is beyond the scope of this
introductory work. Nevertheless, a set of preliminary tests with
different values of $K$ have been conducted. As shown in
Figure~\ref{fig:kimpact}, which refer to the alloca\_test application
and a fixed amount of 500 configurations, initially increasing values
of $K$ seem to improve both pareto (left picture) and objective
distribution (right picture), but after some threshold the $K$ does
not seem to provide any benefit, even returning to slightly worse
results for the last value considered (100). Similar results have been
reported on average for the different benchmarks considered, which are
not included for space reasons. A complete resource with the whole 
results and data obtained in the tests performed can be found
at~\cite{ps_results}.

\begin{table}
	\tbl{Benchmarks\label{tab:defaults}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Default Value} \\
	\hline
	GA population size & 30 \\
	GA crossover probability & 0.8 \\
	GA mutation probability  & 0.1 \\
	$K$ era budget & 30 \\
	alpha threshold & 1 \\
	Total simulations budgets & 100-500 \\
	\hline
	\end{tabular}}
\end{table}

\begin{figure}
  \figLC{kimpact}{Impact of increased K values on Pareto fronts and distribution for the alloca\_test application.}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K.eps} 
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K_ALL.eps} 
  \end{center}
\end{figure}

@@

\begin{figure}
  \figLC{pareto_fronts_500}{Pareto fronts found by PS and GA for a
  fixed budget of 500 configurations.\comment{AA: nelle due figure
  centrali la linea di GA e' continua, mentre in tutte le altre figure
  e' tratteggiata. Il revisore ci chiede di uniformare. Non so come
  riottenere le figure. Potresti pensarci tu?}}
  \begin{center}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
  \end{center}
\end{figure}

Low budget Pareto sets are shown in \figR{pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly dependent upon the application being considered.
These results are obtained in about 5 eras and demonstrate that with a so limited budget PS can compete with
genetic approach while even discovering very different Pareto fronts. 
Considering the extended simulation budget as depicted in
\figR{pareto_fronts_500}, we can qualitatively observe that
even if PS does not strictly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc.
